---
title: "Catholic HAXS"
output:
  word_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

## Intro

This is just a work book from the data analysis that I did with a team at the 2nd Annual CUA data science hackathon. This is mostly meant to show some of my methods when dealing with 11 years of voter data and piecing together a story. Feel free to look over it, this file will soon be refined, as this is the unrefined work of a 5 hour coding sprint. I recomend looking at the end for the draft of the crime explorer app. With some more time invested into it could become a great tool.

## Lets go
Historical data from 1992 to 2018
May need to look at history of all of these elections to get context. We have a couple tasks, find some reasoning behind all these datasets, figure out how to join them together. Or atleast we can join them, into a mega database and then nest them together by year or something. Building a solid database will be the only way to win this.

Starting from 08 the data seems to get tidy.
2010 is when they adopt a tidy data mindset with standardized data.

### Importing

```{r Importing from 2011 to 2018}
library(tidyverse)
library(readr)
imports <- function(){
April_26_2011_Special_Election_Certified_Results <- read_csv("DataElection/DataElection/2011/April_26_2011_Special_Election_Certified_Results.csv")
April_3_2012_Primary_Election_Certified_Results <- read_csv("DataElection/DataElection/2012/April_3_2012_Primary_Election_Certified_Results.csv")
May_15_2012_Special_Election_Certified_Results <- read_csv("DataElection/DataElection/2012/May_15_2012_Special_Election_Certified_Results.csv")
November_6_2012_General_and_Special_Election_Certified_Results <- read_csv("DataElection/DataElection/2012/November_6_2012_General_and_Special_Election_Certified_Results.csv")
April_23_2013_Special_Election_Certified_Results <- read_csv("DataElection/DataElection/2013/April_23_2013_Special_Election_Certified_Results.csv")
April_1_2014_Primary_Election_Certified_Results <- read_csv("DataElection/DataElection/2014/April_1_2014_Primary_Election_Certified_Results.csv")
July_15_2014_Special_Election_Certified_Results <- read_csv("DataElection/DataElection/2014/July_15_2014_Special_Election_Certified_Results.csv")
November_4_2014_General_Election_Certified_Results <- read_csv("DataElection/DataElection/2014/November_4_2014_General_Election_Certified_Results.csv")
April_28_2015_Special_Election_Certified_Results <- read_csv("DataElection/DataElection/2015/April_28_2015_Special_Election_Certified_Results.csv")
June_14_2016_Primary_Election_Certified_Results <- read_csv("DataElection/DataElection/2016/June_14_2016_Primary_Election_Certified_Results.csv")
November_8_2016_General_Election_Certified_Results <- read_csv("DataElection/DataElection/2016/November_8_2016_General_Election_Certified_Results.csv")
December_4_2018_Special_Election_Certified_Results <- read_csv("DataElection/DataElection/2018/December_4_2018_Special_Election_Certified_Results.csv")
June_19_2018_Primary_Election_Certified_Results <- read_csv("DataElection/DataElection/2018/June_19_2018_Primary_Election_Certified_Results.csv")
November_6_2018_General_Election_Certified_Results <- read_csv("DataElection/DataElection/2018/November_6_2018_General_Election_Certified_Results.csv")
bigdata <- list(April_26_2011_Special_Election_Certified_Results, April_3_2012_Primary_Election_Certified_Results, May_15_2012_Special_Election_Certified_Results, November_6_2012_General_and_Special_Election_Certified_Results, April_23_2013_Special_Election_Certified_Results, April_1_2014_Primary_Election_Certified_Results, July_15_2014_Special_Election_Certified_Results, November_4_2014_General_Election_Certified_Results, April_28_2015_Special_Election_Certified_Results, June_14_2016_Primary_Election_Certified_Results, November_8_2016_General_Election_Certified_Results) %>% reduce(rbind)

data2018 <- November_6_2018_General_Election_Certified_Results %>%
  full_join(June_19_2018_Primary_Election_Certified_Results) %>%
  full_join(December_4_2018_Special_Election_Certified_Results)


DF2011to2018 <- data2018 %>%
  full_join(bigdata,  by = c("ElectionDate" = "ELECTION_DATE", "ElectionName" = "ELECTION_NAME", "ContestNumber" = "CONTEST_ID", "ContestName" = "CONTEST_NAME", "PrecinctNumber" = "PRECINCT_NUMBER", "WardNumber" = "WARD", "Candidate" = "CANDIDATE", "Party" = "PARTY", "Votes" = "VOTES"))
return(DF2011to2018)
}
DF2011to2018 <- imports()
```

### EDA

```{r Intro}
library(DataExplorer)
str(DF2011to2018)
plot_intro(DF2011to2018)
plot_missing(DF2011to2018)
plot_bar(DF2011to2018)
plot_qq(DF2011to2018)
#we need to clean ElectionName so it works better
```


```{r Data Clean up}
#Election name to Factor
#Character to Date
# Party to Factor
#Party to Factor
#Precinct and ward to factor, contest to continuous
COPY <- DF2011to2018
COPY$ElectionName <- str_replace(COPY$ElectionName, "D.C. General Election", "General Election")
COPY$ElectionName <- str_replace(COPY$ElectionName, "D.C. Generation Election", "General Election")
COPY$ElectionName <- str_replace(COPY$ElectionName, "D.C. Special Election", "Special Election")
COPY$ElectionName <- str_replace(COPY$ElectionName, "D.C. Primary Election", "Primary Election")
plot_bar(COPY)
```

```{r Date work}
library(lubridate)
COPY$ElectionDate <- str_replace(COPY$ElectionDate, "6/19/2018 12:00:00 AM", "6/19/2018")
COPY$ElectionDate <- str_replace(COPY$ElectionDate, "11/6/2018 12:00:00 AM", "11/6/2018")
COPY$ElectionDate <- str_replace(COPY$ElectionDate, "12/4/2018 12:00:00 AM", "12/4/2018")
COPY$ElectionDate <- as.Date(COPY$ElectionDate, format = "%m/%d/%Y")
COPY$ContestName <- factor(COPY$ContestName)
COPY$WardNumber <- factor(COPY$WardNumber)
COPY$PrecinctNumber <- factor(COPY$PrecinctNumber)
str(COPY)
```

```{r}
# extract totals into a total data frame
unique(COPY$ElectionDate)
COPY2 <- COPY
totals <- COPY2 %>%
  filter(str_detect(ContestName, "REGISTERED")) %>%
  mutate(totals = Votes) %>%
  select(-Votes)

COPY2$ContestName <- tolower(COPY2$ContestName)
removal <- str_detect(COPY2$ContestName, "registered")
COPY2 <- COPY2[-removal, ]

copy3 <- COPY2 %>%
  full_join(totals[, c("ElectionDate", "PrecinctNumber", "WardNumber", "totals")],  by = c("ElectionDate", "PrecinctNumber", "WardNumber"))
```

```{r}
library(readxl)
datelist <- read_excel("datelist.xlsx")
str(datelist)
COPY2$ElectionDate <- as.POSIXct(COPY2$ElectionDate)
COPY2 %>%
  left_join(datelist,  by = c("ElectionDate" = "Date")) %>%
  ggplot(aes(x = ElectionDate, y = Votes, fill = AvgTemp)) + 
  geom_bar(stat = "identity")
```
```{r}
totals2 <- totals %>%
  group_by(WardNumber, ElectionDate) %>%
  summarize(totalsum = sum(totals))
COPY2 %>%
  filter(str_detect(ContestName, "total")) %>%
  group_by(WardNumber, ElectionDate) %>%
  summarize(Votesum = sum(Votes)) %>%
  ggplot(aes(x = WardNumber, y = Votesum)) + 
  geom_bar(stat = "identity", fill = "green") + 
  geom_bar(data = totals2, aes(x = WardNumber, y = totalsum), fill = "red", alpha = 0.8, stat = "identity", inherit.aes = FALSE) 
```


```{r}
library(ggthemes)
COPY2 %>%
  group_by(WardNumber) %>%
  summarize(Votesum = sum(Votes)) %>%
  ggplot(aes(x = WardNumber, y = Votesum)) +
  geom_bar(stat = "identity") + 
  theme_stata() + 
  xlab("Ward") + 
  ylab("Sum of Votes") +
  ggtitle("Votes per Ward", subtitle = "2011 to 2018")
```
```{r}
import_crime <- function(){
Crime_Incidents_in_2013 <- read_csv("drive-download-20190330T183559Z-001/Crime_Incidents_in_2013.csv")
Crime_Incidents_in_2014 <- read_csv("drive-download-20190330T183559Z-001/Crime_Incidents_in_2014.csv")
Crime_Incidents_in_2015 <- read_csv("drive-download-20190330T183559Z-001/Crime_Incidents_in_2015.csv")
Crime_Incidents_in_2016 <- read_csv("drive-download-20190330T183559Z-001/Crime_Incidents_in_2016.csv")
Crime_Incidents_in_2017 <- read_csv("drive-download-20190330T183559Z-001/Crime_Incidents_in_2017.csv")
Crime_Incidents_in_2018 <- read_csv("drive-download-20190330T183559Z-001/Crime_Incidents_in_2018.csv")
crime_data <- list(Crime_Incidents_in_2013, Crime_Incidents_in_2014, Crime_Incidents_in_2015, Crime_Incidents_in_2016, Crime_Incidents_in_2017, Crime_Incidents_in_2018) %>% reduce(rbind)
return(crime_data)
}
crime_data <- import_crime()
crime_sum <- crime_data %>%
  group_by(WARD) %>%
  summarize(n = n()) %>%
  na.omit()
crime_data %>%
  group_by(OFFENSE) %>%
  summarise()
crime_data %>%
  filter(OFFENSE == "HOMICIDE") %>%
  group_by(WARD) %>%
  summarize(crimenumber = n()) %>%
  ggplot(aes(x = factor(WARD), y = crimenumber)) + 
  geom_bar(stat = "identity") +
  xlab("Ward") +
  theme_stata() +
  ylab("Frequency of Homicide") +
  ggtitle("Homicide frequency per ward", subtitle = "2013 to 2018")
```

## Spatial Visualizations

```{r}
library(ggmap)
DCMAP <- get_map("Washington, DC", zoom = 12)
set.seed(20)
clusters <- kmeans(crime_data[ ,1:2], 8)
crime_data$ESTCLUSTER <- as.factor(clusters$cluster)
ggmap(DCMAP) + geom_point(data = crime_data, aes(x = X, y = Y, col = factor(ESTCLUSTER)), alpha = 0.4) + 
  ggtitle("DC Wards using Kmeans on crime data")
ggmap(DCMAP) + geom_point(data = crime_data, aes(x = X, y = Y, col = factor(WARD)), alpha = 0.4) 
ggmap(DCMAP) + 
  geom_bin2d(data = crime_data %>% filter(OFFENSE == 'HOMICIDE'), aes(x = X, y = Y), alpha  = 0.8)
```

```{r}
library(tidyverse)
library(maptools)
library(sp)
library(mapdata)
library(rgl)
library(rgdal)
library(ggplot2)
library(rgeos)
precinct <- readOGR("C:/Users/bingo/Desktop/CUA HaXS/Voting_Precinct__2012/Voting_Precinct__2012.shp") 
ward <- readOGR("C:/Users/bingo/Desktop/CUA HaXS/Ward_from_2012/Ward_from_2012.shp")
precinct2 <- fortify(precinct)
ward2 <- fortify(ward)
precinctdat <- as.data.frame(precinct)
map1 <- ggplot() + 
  geom_polygon(data = precinct2, aes(long, lat, group = group), 
               colour = "black", fill = "white") +
  geom_polygon(data = ward2, aes(long, lat, group = group, fill = group),
               colour = "black", alpha = 0.3, size = 1) +
  ggtitle("DC Precincts and Wards",
          subtitle = "Est. 2012") +
  coord_quickmap()+
  theme_minimal()
map1
#opens up possibility of DC chloropleth map (interactive too)
library(ggmap)
library(sf)
DCMAP <- get_map("Washington, DC", zoom = 12)
wardsp <- spTransform(ward, CRS("+proj=longlat +datum=WGS84"))
wardsp2 <- fortify(wardsp)
#Add another column for ward
WARDWITHLABELS <- ward2 %>%
  mutate(WARD = 
           case_when(
             group == 0.1 ~ 8,
             group == 1.1 ~ 6,
             group == 2.1 ~ 7,
             group == 3.1 ~ 2,
             group == 4.1 ~ 1, 
             group == 5.1 ~ 5,
             group == 6.1 ~ 3,
             group == 7.1 ~ 4
           ))

# Get centroids to plot labels
ward_Cent <- gCentroid(ward, byid = TRUE)
ward_CentPoints <- as.data.frame(ward_Cent)
Wardlabs <- ward_CentPoints %>%
  cbind(c(8, 6, 7, 2, 1, 5, 3, 4))
colnames(Wardlabs) <- c("x", "y", "WARD")
# Plot the points on the wards
ggplot() +
  geom_polygon(data = WARDWITHLABELS, aes(long, lat, group = group, fill = factor(WARD)),
               colour = "black", alpha = 0.3, size = 1) +
  geom_text(data = Wardlabs, aes(x, y, label = WARD), size = 5) +
  geom_point(data = crime_data %>% filter(OFFENSE == "HOMICIDE"), aes(x = X, y = Y), alpha = 0.4) +
  ggtitle("DC Wards",
          subtitle = "Est. 2012") +
  theme(legend.position = "none") +
  coord_quickmap()
```
```{r}
#experiment with google maps
ggmap(DCMAP) + 
  geom_polygon(data = WARDWITHLABELS, aes(long, lat, group = group, fill = factor(WARD)),
               colour = "black", alpha = 0.3, size = 1)
library(plotly)
Sys.setenv('MAPBOX_TOKEN' = 'pk.eyJ1IjoibWFzdGVyYmluZ28xIiwiYSI6ImNqdDluOHo2aDAxenQ0OW51dmdkOGIyaDkifQ.KMv1Wkds1VtmtzOOmMWuiw')
plot_mapbox(mode = 'scattermapbox') %>%
  add_polygons(data = WARDWITHLABELS , 
            x = ~ long, y = ~ lat, split = ~factor(WARD), 
            text = ~WARD, hoverinfo = "none", alpha = 0.3) %>%
  add_markers(data = crime_data %>% filter(OFFENSE == "HOMICIDE"), x = ~X, y = ~Y, alpha = 1, color = "blue") %>%
  layout(mapbox = list(zoom = 10.5,
                       center = list(lat = ~median(WARDWITHLABELS$lat),
                                     lon = ~median(WARDWITHLABELS$long))
  ))
```

Attempting to make a small web app out of the crime data with a explorer.

```{r}
library(crosstalk)
dooot <- SharedData$new(crime_data %>% filter(OFFENSE == "HOMICIDE") %>%
                          select(X, Y, METHOD, WARD))
bscols(
  plot_mapbox() %>%
    add_markers(data = dooot, x = ~X, y = ~Y, text = ~WARD, hoverinfo = "text") %>%
    highlight(on = "plotly_selected", dynamic = TRUE) %>%
    layout(mapbox = list(zoom = 10.5,
                       center = list(lat = ~median(WARDWITHLABELS$lat),
                                     lon = ~median(WARDWITHLABELS$long)))), 
  DT::datatable(dooot)
)
```


